# `leaflet` and `mapview` maps {#leaflet}

```{r setup070, include=FALSE}
# path to this file name
if (!interactive()) {
    fnamepath <- current_input(dir = TRUE)
} else {
    fnamepath <- ""
}
```

This exercise will introduce construction of a Leaflet map using the `leaflet` package. Also simple interface uses the `mapview::mapview()` function, which automatically adds the OpenStreetMap background, uses pre-loaded colors, and with the `legend = TRUE` option also adds a map legend.

```{r}
mapview(nhood_pov, zcol = "pct_pov", legend = TRUE)
```

We will now build up a map using `leaflet()` for a bit more control.

```{r}
# CRS
nhood_pov_4326 <- nhood_pov %>% st_transform(4326)

# make a label with the count of persons, persons below poverty, and % poverty
nhood_pov_4326$mylab <- sprintf("n_pov=%s<br>n=%s<br>%s<br>", 
                                round(nhood_pov_4326$n_pov, 0), 
                                round(nhood_pov_4326$n, 0), 
                                paste("pov=", nhood_pov_4326$pct_pov, "%", sep=""))

l <- leaflet(data = nhood_pov_4326) %>%
    addPolygons(popup = ~mylab, weight = 2) %>%
    addTiles()
l
```

Let's change the labels to include neighborhood name and % poverty, add choropleth fill and a legend:

```{r}
# CRS
nhood_pov_4326 <- nhood_pov %>% st_transform(4326)

mypalette <- colorQuantile(palette = "viridis", domain = nhood_pov_4326$pct_pov, n = 4)

# make a label with the count of persons, persons below poverty, and % poverty
nhood_pov_4326$mylab <- sprintf("%s<br>%s<br>", 
                                nhood_pov_4326$gen_alias, 
                                paste("pov=", nhood_pov_4326$pct_pov, "%", sep=""))

l <- leaflet(data = nhood_pov_4326) %>%
    addPolygons(popup = ~mylab, 
                weight = 2, 
                fillColor = ~mypalette(pct_pov),
                opacity = 0.7) %>%
    addTiles() %>%
    addLegend(pal=mypalette, values=~pct_pov, opacity=0.7, title = "% below poverty", position = "bottomleft" )

l
```

Finally, add the hospitals with a label that will display when hovering over the point marker.

```{r}
# CRS
nhood_pov_4326 <- nhood_pov %>% st_transform(4326)

# hospitals
hospitals <- st_read(file.path(mydatadir, "medical_facilities", "medical_facilities.shp"), quiet = TRUE) %>% st_transform(4326)

mypalette <- colorQuantile(palette = "viridis", domain = nhood_pov_4326$pct_pov, n = 4)

# make a label with the count of persons, persons below poverty, and % poverty
nhood_pov_4326$mylab <- sprintf("%s<br>%s<br>", 
                                nhood_pov_4326$gen_alias, 
                                paste("pov=", nhood_pov_4326$pct_pov, "%", sep=""))

l <- leaflet(data = nhood_pov_4326) %>%
    addPolygons(popup = ~mylab, 
                weight = 2, 
                fillColor = ~mypalette(pct_pov),
                opacity = 0.7) %>%
    addTiles() %>%
    addLegend(pal=mypalette, values=~pct_pov, opacity=0.7, title = "% below poverty", position = "bottomleft" )

l <- addCircleMarkers(map = l, 
                     data = hospitals,
                     radius = 5, 
                     weight = 1,
                     opacity = 0.9,
                     fillOpacity = 0.5,
                     label = ~ABB_NAME)

l
```
